Author: Justin Mau
LRES 525



Libraries:
```{r}
library(ggplot2)
library(neonUtilities)
library(geoNEON)
library(tidyverse)
library(terra)
#library(rhdf5)
library(httr)
library(jsonlite)
library(downloader)
library(ggplot2)
library(RStoolbox)
library(sp)
library(raster)
library(neonOS)
#library(rgdal)
library(dplyr)
```


Downloading the Vegetation Structure vst product for the UNDE site in 2020-08
```{r}
vst <- loadByProduct(dpID="DP1.10098.001", 
                     site="UNDE",
                     startdate="2020-08",
                     enddate="2020-09", 
                     check.size=F)
```



Reads in lat/long and such for the specific plot I'm looking at
```{r}
#targetPlot = "UNDE_037"
targetPlot = "UNDE_065"
targetPlot_info = vst$vst_perplotperyear[which(vst$vst_perplotperyear$plotID==targetPlot),]
targetPlot_lat = targetPlot_info$decimalLatitude
targetPlot_lon = targetPlot_info$decimalLongitude
targetPlot_northing = targetPlot_info$northing
targetPlot_easting = targetPlot_info$easting
```




Loading in all apparent individual trees for the specified plot
```{r}
vst.trees <- vst$vst_apparentindividual[which
                                        (vst$vst_apparentindividual$plotID== targetPlot),]

# creates one table with the known individuals and their relevant information
vst.trees <- merge(vst.trees, vst$vst_mappingandtagging, by = "individualID")
vst.trees <- dplyr::select(vst.trees, uid.x, individualID, scientificName, taxonID, growthForm, plantStatus, canopyPosition, plotID.x)

View(vst.trees)
```




generating simple plots about distribution -------------------------------------------------------------------------
--------------------------------------------------------------------------------------------------------------------
Plot for tree counts by species
```{r}
tree_counts <- vst.trees %>% count(scientificName)

ggplot(tree_counts, aes(x = scientificName, y = n)) +
  geom_bar(stat = "identity") +
  labs(x = "Tree Species", y = "Count", title = "Distribution of Tree Species") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```

Plot for tree counts by plant status
```{r}
tree_counts <- vst.trees %>% count(plantStatus)

ggplot(tree_counts, aes(x = plantStatus, y = n)) +
  geom_bar(stat = "identity") +
  labs(x = "Tree Status", y = "Count", title = "Distribution of Tree Status") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```

Plot for tree counts by canopy position (!)
```{r}
tree_counts <- vst.trees %>% count(canopyPosition)

ggplot(tree_counts, aes(x = canopyPosition, y = n)) +
  geom_bar(stat = "identity") +
  labs(x = "Canopy Position", y = "Count", title = "Distribution of Tree Canopy Positions") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```

In plot 065, there are 32 plants in full sun and 25 only partially shaded
So there should be at least 32 trees identified in the canopy
Full model only identified about 20
```{r}
length(vst.trees$canopyPosition[which(vst.trees$canopyPosition == "Full sun")])
length(vst.trees$canopyPosition[which(vst.trees$canopyPosition == "Partially shaded")])
```
-------------------------------------------------------------------------------



Loading in tree location data -------------------------------------------------
-----------------------------------------------------------------------------
```{r}
vst.loc <- getLocTOS(data=vst$vst_mappingandtagging, dataProd="vst_mappingandtagging")
```


```{r}
# uses neonOS to join all these tables together
veg <- joinTableNEON(vst.loc, 
                     vst$vst_apparentindividual,
                     name1="vst_mappingandtagging",
                     name2="vst_apparentindividual")

# filters out NA values
veg <- veg[!is.na(veg$adjDecimalLongitude) & !is.na(veg$adjDecimalLongitude), ]

```


For just extracting for plot065, this code should be run
```{r}
# filters to just specific plot and canopy position
veg <- veg[which(veg$plotID == "UNDE_065"), ]
veg <- veg[which(veg$canopyPosition == "Full sun" | veg$canopyPosition == "Partially shaded"), ]

x <- veg$adjDecimalLongitude
y <- veg$adjDecimalLatitude
species <- veg$scientificName
uncertainty <- veg$adjCoordinateUncertainty
canopyPosition <- veg$canopyPosition
crownDiameter <- veg$maxCrownDiameter

# creates spatial point to put on the plot
tree_points <- SpatialPoints(coords = cbind(x, y), proj4string = CRS("+proj=longlat +datum=WGS84"))
tree_points$species <- species
tree_points$plotID <- "plot065"
tree_points$uncertainty <- uncertainty
tree_points$canopyPosition <- canopyPosition
tree_points$crownDiameter <- crownDiameter


# Convert spatial points to data frame
tree_data <- as.data.frame(tree_points)
# Export to CSV
write.csv(tree_data, paste0("./Tree_Location_Points/tree_locations_", "plot065", ".csv"), row.names = FALSE)

```

OR 

For extracting for ALL plots, this code should be run instead
```{r}

for (tp in unique(veg$plotID)) {
  
  x <- veg$adjDecimalLongitude[which(veg$plotID == tp)]
  y <- veg$adjDecimalLatitude[which(veg$plotID == tp)]
  species <- veg$scientificName[which(veg$plotID == tp)]
  uncertainty <- veg$adjCoordinateUncertainty[which(veg$plotID == tp)]
  
  # creates spatial point to put on the plot
  tree_points <- SpatialPoints(coords = cbind(x, y), proj4string = CRS("+proj=longlat +datum=WGS84"))
  tree_points$species <- species
  tree_points$plotID <- tp
  tree_points$uncertainty <- uncertainty
  
  
  # Convert spatial points to data frame
  tree_data <- as.data.frame(tree_points)
  # Export to CSV
  write.csv(tree_data, paste0("./Tree_Location_Points/tree_locations_", tp, ".csv"), row.names = FALSE)
  
}
```
-------------------------------------------------------------------------------











-----------------------------------------------------------------------------
Now using the code from lab 5 to load in spectroscopy data for UNDE site
```{r}
req.aop <- GET("http://data.neonscience.org/api/v0/products/DP3.30010.001")
avail.aop <- fromJSON(content(req.aop, as="text"), simplifyDataFrame=T, flatten=T)

## note: use round parenthesis to create AND display new variable in the console in one step
spec.urls <- unlist(avail.aop$data$siteCodes$availableDataUrls)

#2020-08
sel <- GET(spec.urls[intersect(grep("UNDE", spec.urls),grep("2020-08", spec.urls))])

sel.files <- fromJSON(content(sel, as="text"))
sel.files$data$files$name
```

Now finds out the easting/northing values, mostly to figure out which tiles to choose
```{r}
print(targetPlot_easting)
print(targetPlot_northing)
h <- sel.files$data$files$name[(grep("30[3-4]000_512[2-3]000", sel.files$data$files$name))]
h <- h[3]
```

Downloads the tile(s) that contains the plot
```{r}
download(url = sel.files$data$files$url[grep(h, sel.files$data$files$name)],
         paste0("./",h), mode="wb")
```

Load it in to R to view
```{r}
hires <- rast("./mergedRaster.tif")
plotRGB(hires, stretch="lin")
```



Getting the LAI Spectrometer imagery
(Unfortunately LAI wouldn't work with the model since the model needs 3 bands)
```{r}
req.aop <- GET("http://data.neonscience.org/api/v0/products/DP1.00033.001")
avail.aop <- fromJSON(content(req.aop, as="text"), simplifyDataFrame=T, flatten=T)

## note: use round parenthesis to create AND display new variable in the console in one step
spec.urls <- unlist(avail.aop$data$siteCodes$availableDataUrls)

#2020-08
sel <- GET(spec.urls[intersect(grep("UNDE", spec.urls),grep("2020-08", spec.urls))])

sel.files <- fromJSON(content(sel, as="text"))
sel.files$data$files$name

h <- sel.files$data$files$name[(grep("304000_5123000", sel.files$data$files$name))]
h
h <- h[1]
#download(url = sel.files$data$files$url[grep(h, sel.files$data$files$name)],
#         paste0("./",h), mode="wb")

download(url = "https://phenocam.nau.edu/webcam/sites/NEON.D05.UNDE.DP1.00033/",
         "./", mode="wb")
```





